Online Verification of Commutativity

### 2020-06-17 acmPaper/paper.tex

In order to verify commutativity efficiently over online addition, we identified two key insights.
The first was that in a commutative diagram, when a new edge is to be added, at most one path needs to be checked against for a given source and sink pair.
Because the diagram commutes, all the paths between a given source and sink are equal and a representative to check against can arbitrarily be chosen.
It lead to an $O(|V|^2(|E|+|V|)$ algorithm to verify a diagram remains commutative over the course of online addition, assuming an oracle to check the equality of functions.
The algorithm makes an asymptotically optimal number of calls to the oracle.

@@

We allowed an "error tolerance" for floating point error.
Graphs were generally found to be consistent up to an error tolerance to the order of magnitude $10^{-6}$.
However, below this threshold we found inconsistencies.
Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using algorithm ~\ref{optimal}, and in 133$\pm$13 seconds with algorithm ~\ref{polynomial}. 
Computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
More information about implementation can be found in the evaluation section.

>>>>

In the course of efficiently verifying commutativity over online addition, we identified two key insights.
The first was that in a commutative diagram, when a new edge is to be added, at most one path needs to be checked against for a given source and sink pair.
Because the diagram commutes, all the paths between a given source and sink are equal and a representative to check against can arbitrarily be chosen.
It lead to an $O(|V|^2(|E|+|V|)$ algorithm to verify a diagram remains commutative over the course of online addition, assuming an oracle to check the equality of functions.
The algorithm makes an asymptotically optimal number of calls to the oracle.

@@

We allowed an "error tolerance", so that differences reported would not be the trivial consequences of a floating point error.
However, this relaxation of the equality oracle into imprecision meant that the mathematical reasoning that allow the algorithms to remove redundant path checks no longer applied.
For instance, composing a new function with two approximately equal functions does not lead to equal results, so theorem ~\ref{reductionRule} fails with this approximate equality.
When the algorithms reported no inconsistencies, it was still possible that the graph possessed inconsistencies above the given threshold and did not commute.
Nonetheless, both algorithms were effective in catching inconsistencies. Algorithm ~\ref{polynomial} started finding inconsistencies at error tolerances to the order of $10^{-3}$, and ~\ref{optimal}, which makes more invalid redundant path check removals, at error tolerances to the order of $10^{-7}$.

Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using algorithm ~\ref{optimal}, and in 133$\pm$13 seconds with algorithm ~\ref{polynomial}. 
Computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
More information about implementation can be found in the evaluation section.
